---
title: "SCALE Report"
author: "Montana"
date: "2024-02-09"
output: html_document
---

```{r, echo = F, message = F, warning = F}

setwd("C:/Users/monta/OneDrive - Airey Family/GitHub/AFRP/MA2276_Code/SCALE")
library(dplyr)
library(tidyr)
library(tidyverse)
library(gridExtra)
`%nin%` = Negate(`%in%`)
## Month setup
month_bin = c(0,7,12,15)
## Data setup
s.d = read.csv(file = "SCALE_MTS_2023.csv") %>% separate(YSAMP, into = c("GEAR", "WATER", "YEAR", "YSAMP_N"), remove = F)  %>% 
  mutate(FAMILY = tolower(FAMILY)) %>%
  mutate(ORDER..OR.ABOVE. = tolower(ORDER..OR.ABOVE.)) %>%## Scale Data 
  filter(ORDER..OR.ABOVE. != "") %>%
  mutate(FAMILY = str_replace(FAMILY, " ", "")) %>%
  mutate(ORDER..OR.ABOVE. = str_replace(ORDER..OR.ABOVE., " ", ""))

sample = read.csv(file = "SCALE_sample.csv") %>%
  mutate(YEAR = as.character(YEAR))

## How many lakes are each common family found in 
common = s.d %>% 
  filter(ORDER..OR.ABOVE. != "") %>%
  filter(FAMILY != "unidentified") %>% 
  filter(FAMILY != "terrestrial") %>%
  filter(FAMILY != "") %>%
  filter(FAMILY != " ") %>%
  filter(FAMILY != "degraded") %>%
  group_by(WATER, ORDER..OR.ABOVE., FAMILY) %>% 
  summarize(sum = sum(TOTAL_N)) %>% 
  mutate(FAMILY = tolower(FAMILY)) %>%
  mutate(ORDER..OR.ABOVE. = tolower(ORDER..OR.ABOVE.)) %>% 
  filter(ORDER..OR.ABOVE. != "") %>%
  ungroup() %>%
  group_by(ORDER..OR.ABOVE., FAMILY) %>%
  summarize(total_lakes = n()) 


full = left_join(s.d, sample, by = c("YSAMP"))  %>%
  mutate(MONTH = .bincode(MONTH, month_bin)) %>%
  mutate(MONTH = as.character(MONTH)) %>%
  mutate(MONTH = str_replace(MONTH, "2", "FALL")) %>%
  mutate(MONTH = str_replace(MONTH, "1", "SPRING")) %>%
  select(MONTH, YSAMP, GEAR.x,SITE_N, WATER.x, YEAR.x, YSAMP_N, MTS_N, GROUP, ORDER..OR.ABOVE., FAMILY, GENUS, SPECIES, TOTAL_N) %>%
  rename("GEAR" = GEAR.x, "WATER" = WATER.x, "YEAR" = YEAR.x)

save(file = "FullData.RData", full)

common_month = full %>% 
  filter(ORDER..OR.ABOVE. != "") %>%
  filter(FAMILY != "unidentified") %>% 
  filter(FAMILY != "terrestrial") %>%
  filter(FAMILY != "") %>%
  filter(FAMILY != " ") %>%
  filter(FAMILY != "degraded") %>%
  group_by(WATER,ORDER..OR.ABOVE., FAMILY, MONTH, SITE_N) %>% 
  summarize(sum = sum(TOTAL_N)) %>% 
  mutate(FAMILY = tolower(FAMILY)) %>%
  mutate(ORDER..OR.ABOVE. = tolower(ORDER..OR.ABOVE.)) %>% 
  filter(ORDER..OR.ABOVE. != "")
```

#### Distribution of macroinvertebrate families across lakes

Libellulids are found across 11 of 12 lakes

```{r,echo = F, message = F, warning = F}

s.d %>% 
  left_join(common) %>%
  
  group_by(WATER, ORDER..OR.ABOVE., FAMILY, total_lakes) %>% 
  summarize(sum = sum(TOTAL_N)) %>% 
  ungroup() %>% 
  mutate(FAMILY = tolower(FAMILY)) %>%
  mutate(ORDER..OR.ABOVE. = tolower(ORDER..OR.ABOVE.)) %>% 
  filter(ORDER..OR.ABOVE. != "") %>% 
  arrange(ORDER..OR.ABOVE., ORDER..OR.ABOVE.)  %>%
  filter(total_lakes != "NA")%>% 
  
  ggplot(aes(y = FAMILY, x = sum, fill = WATER)) + 
  geom_bar(stat = 'identity', position = 'dodge') + 
  theme(axis.text.x = element_text(angle = 90)) + 
  facet_wrap(~total_lakes, scales = "free")



```

#### Seasonal Differences within lake

Only 8 lakes were sampled twice. The following graph shows all taxa that were sampled during both seasons at each lake.
```{r,echo = F, message = F, warning = F}


count_month = common_month %>%
  filter(MONTH != "NA") %>%
  unique() %>%
  mutate(sum = as.numeric(sum)) %>%
  ungroup() %>%
  group_by(WATER, ORDER..OR.ABOVE., MONTH, FAMILY) %>%
  summarize(total_inverts = sum(sum)) %>%
  pivot_wider(names_from = MONTH, values_from = total_inverts) %>%
  filter(is.na(FALL) != T & is.na(SPRING)!= T) %>%
  ungroup() %>%
  group_by(FAMILY) %>%
  summarize(total_lakes = n())

common_month %>%
  filter(MONTH != "NA") %>%
  unique() %>%
  mutate(sum = as.numeric(sum)) %>%
  ungroup() %>%
  group_by(WATER, ORDER..OR.ABOVE., MONTH, FAMILY) %>%
  summarize(total_inverts = sum(sum)) %>%
  pivot_wider(names_from = MONTH, values_from = total_inverts) %>%
  filter(is.na(FALL) != T & is.na(SPRING)!= T) %>%
  left_join(count_month) %>%
  ggplot(aes(x = reorder(FAMILY, -total_lakes), fill = WATER)) + 
  geom_bar() + 
  theme_minimal() + 
  labs(fill = "Lake") + 
  theme(axis.text.x = element_text(angle = 90, vjust = .5)) + 
  ylab("Number of Lakes") + 
  xlab("Family")
```

#### Site differences within lakes 

We sampled two lakes (Dart and Moss Lake) at high site (N = 4) intensity in the 2023 spring


```{r,echo = F, message = F, warning = F}

msl = full %>% filter(WATER %in% c("MSL"), 
                MONTH == "SPRING") %>%
  group_by(SITE_N, WATER, FAMILY) %>%
  summarize(total_interts = sum(TOTAL_N)) %>%
  filter(FAMILY != "degraded") %>%
  ungroup() %>%
  group_by(FAMILY) %>%
  summarize(sum_sites = n()) %>%
  ggplot(aes(x = reorder(FAMILY, -sum_sites), y = sum_sites)) + 
  geom_bar(stat = "identity") +
  ylab("Total Sites Present") + 
  xlab("Family") + 
  theme_minimal() + 
  theme(axis.text.x = element_text(angle = 90, vjust = .5)) + 
  ggtitle("Moss Lake")


dtl = full %>% filter(WATER %in% c("DTL"), 
                MONTH == "SPRING") %>%
  group_by(SITE_N, WATER, FAMILY) %>%
  summarize(total_interts = sum(TOTAL_N)) %>%
  filter(FAMILY != "degraded") %>%
  ungroup() %>%
  group_by(FAMILY) %>%
  summarize(sum_sites = n()) %>%
  ggplot(aes(x = reorder(FAMILY, -sum_sites), y = sum_sites)) + 
  geom_bar(stat = "identity") +
  ylab("Total Sites Present") + 
  xlab("Family") + 
  theme_minimal() + 
  theme(axis.text.x = element_text(angle = 90, vjust = .5)) + 
  ggtitle("Dart Lake")

grid.arrange(msl, dtl, ncol = 2)


```

#### The other lakes were not intensively sampled across space, but were sampled at two time points

```{r, out.height="200%", out.width="100%",echo = F, message = F, warning = F}


full %>% filter(WATER %nin% c("DTL","MSL", "COM", "LML","GNL","ETL"), MONTH != "NA") %>%
  group_by(SITE_N, WATER, MONTH,FAMILY) %>%
  summarize(total_interts = sum(TOTAL_N)) %>%
  filter(FAMILY != "degraded") %>%
  ungroup() %>%
  group_by(WATER,FAMILY, MONTH) %>%
  summarize(sum_sites = n()) %>%
  ggplot(aes(x = FAMILY, y = sum_sites, fill = MONTH)) +
  geom_bar(stat = 'identity') +
  facet_wrap(~WATER, scales = "free") +  ylab("Total Sites Present") + 
  xlab("Family") + 
  theme_minimal() + 
  theme(axis.text.x = element_text(angle = 90, vjust = .5, size = 6)) + 
  ggtitle("Dart Lake")

```

##### Short-list of taxa that are found seasonally and are common across both regional and within-lake spatial scales

The below taxa are found at 100% of the sites within their lakes and are observed in both spring + fall sampling in 50% of the seasonally sampled lakes.

```{r,echo = F, message = F, warning = F}

num_sites_sampled = full %>% select(WATER, MONTH, SITE_N) %>% unique() %>%
  group_by(WATER, MONTH) %>%
  summarize(total.sites = n())

full %>% 
  left_join(num_sites_sampled) %>%
  group_by(SITE_N, WATER, MONTH,FAMILY, total.sites) %>%
  summarize(total_interts = sum(TOTAL_N)) %>%
  filter(FAMILY != "degraded") %>%
  ungroup() %>%
  group_by(WATER,FAMILY, MONTH, total.sites) %>%
  summarize(sum_sites = n()) %>%
  mutate(percent_sites = round(sum_sites / total.sites * 100, digits = 0)) %>%
  filter(percent_sites ==100) %>% 
  filter(FAMILY %in% (count_month %>% filter(total_lakes > 4))$FAMILY) %>%
  ungroup() %>%
  select(FAMILY) %>% unique()


```


##### Occurance of mollusk family across sampled lakes

```{r,echo = F, message = F, warning = F}

mullusks = full %>% filter(ORDER..OR.ABOVE. %in% c("bivalvia","gastropoda")) 

mullusks %>% select(FAMILY, WATER) %>% group_by(FAMILY) %>% 
  unique() %>%
  summarize(Total_Lakes = n()) %>%
  mutate(Percent_Lakes = round(Total_Lakes / 12*100), digits = 0) %>% 
  select(FAMILY, Percent_Lakes)


snails = mullusks %>% filter(FAMILY %in% c("planorbiae", "lymnaeidae", "viviparidae", "bithyniidae"))

clams = mullusks %>% filter(FAMILY %in% c("unioniae", "spaeriidae"))

```

##### Occurance of mollusk order across sampled lakes

```{r,echo = F, message = F, warning = F}

## Total number of lakes observed in

mullusks %>% select(ORDER..OR.ABOVE., WATER) %>% unique() %>% group_by(ORDER..OR.ABOVE.) %>%
  summarize(Percent_Lakes = round(n() / 12 * 100, digits = 0))
```


```{r}

compiled = read.csv("SCALE_compiled.csv") %>% 
  filter(MONTH != "NA") %>%
  separate(YSAMP, into = c("WATER", "MONTH","YEAR")) %>%
  mutate(MONTH = as.numeric(MONTH)) %>%
  mutate(MONTH = case_when(MONTH < 7 ~ "SPRING", MONTH > 7 ~ "FALL" )) %>%
  rename(SAMPLE = SAMP_N) %>%
  select(WATER, MONTH,SAMPLE, GROUP)

compiled

x = full %>% select(MONTH, WATER,MTS_N) %>% 
  filter(MTS_N %nin% c("NULL.DEGRADED", "NULL")) %>%
  rename(SAMPLE = MTS_N)%>%
  mutate(GROUP = "INVERT") %>%
  rbind(compiled) %>%
  group_by(MONTH, WATER, GROUP) %>%
  summarize(count = n()) %>%
  pivot_wider(names_from = GROUP, values_from = count) %>%
  arrange(WATER) %>%
  filter(INVERT != "NA") 

write.csv(x, "sample_log.csv")
 
dev.off()


full %>%
  group_by(MONTH) %>%
  summarize(count = n() )




```

## Sagamore
```{r}

s.d %>% filter(WATER == "SEL") %>%
  select(ORDER..OR.ABOVE., FAMILY, GENUS) %>%
  unique()
  
```
## Choosing HS samples
```{r}

hs = read.csv("HS_samples.csv")
library(dplyr)

hs_inverts = hs %>% filter(GROUP == "MACROINVERTEBRATES", 
              LOADED.FOR.ISOTOPE..Y.N. == "Y",
              LEFTOVER.SAMPLE..0.5. > 2) %>%
  rename(MTS_N = SAMP_N)

invert_id = read.csv("SCALE_MTS_2023.csv") %>% select(YSAMP, MTS_N, ORDER..OR.ABOVE., FAMILY) %>%
  filter(MTS_N != "")

hs_inverts %>% left_join(invert_id)

scale_sample = read.csv("SCALE_sample.csv")

dat = left_join(hs_inverts, scale_sample) %>%left_join(invert_id)

sum = dat %>% group_by(WATER, MONTH, ORDER..OR.ABOVE.) %>%
  summarize(total = n()) %>%
  filter(total > 3)
sum

full = dat %>%  group_by(WATER, MONTH, ORDER..OR.ABOVE.) %>%
      mutate(total = n()) 


full = full %>% 
  select(YSAMP, MTS_N, GROUP, LEFTOVER.SAMPLE..0.5., WATER, MONTH, ORDER..OR.ABOVE., FAMILY ) 

write.csv(full,"HS_selection.csv")


sum$WATER %>% unique() %>% length()
```

```{r}
library(tidyverse)
library(tidyr)
fish = hs %>% filter(GROUP == "FISH", 
              LOADED.FOR.ISOTOPE..Y.N. == "Y",
              LEFTOVER.SAMPLE..0.5. > 3) %>%
  rename(MTS_N = SAMP_N)


fish_id = read.csv("SCALE_MTS_2023.csv") %>% select(YSAMP, MTS_N, ORDER..OR.ABOVE., FAMILY) %>%
  filter(MTS_N != "")


fish_selection = fish %>% separate(YSAMP, into = c("WATER", "MONTH","YEAR")) %>%
  group_by(MONTH, WATER) %>%
  filter(LEFTOVER.SAMPLE..0.5. > 2) %>%
  arrange(MONTH, WATER, desc(LEFTOVER.SAMPLE..0.5.)) 

write.csv(fish_selection, "fish_selection.csv")


```

